---
title: "July"
author: "Andrea Odell"
date: "6/30/2021"
output: pdf_document
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE)
```

### July 1, 2021

This week, I did a literature search on "Conditional age-at-length" (CAAL) and was pretty unsuccessful. Much of the literature spoke about integrated stock assessments and estimating CAAL within the stock assessment as opposed to estimating through data externally and using those fixed estimates in stock assessments. I felt like what I was finding wasn't relevant to what I needed since I'm just trying to estimate the growth curve parameters to predict missing ages in the dataset. During my search, however, I did find a paper that uses Bayesian length-at-age models to estimate the growth curve parameters. So, I think I'll give it a shot...

Notes. 


```{r cars}
summary(cars)
```

## Including Plots

You can also embed plots, for example:

```{r pressure, echo=FALSE}
plot(pressure)
```

Note that the `echo = FALSE` parameter was added to the code chunk to prevent printing of the R code that generated the plot.
