---
title: "Meeting Notes"
author: "Andrea Odell"
date: "6/24/2021"
output: pdf_document
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE)
knitr::opts_knit$set(root.dir = "/Users/andrea/Box/UC DAVIS/Oken Lab/lingcod-rockfish")
```

### June 24 2021

This week I went back to the lingcod dataset collected by Bonnie et al. and went through the data cleaning script "clean.R". Updated a few things to make the dataset more tidy. Created new total Sebastes by number and by weight columns by summing all sebastes species into one total sebastes (X..Sebastes.Total and wt..Sebastes.Total). Then, I updated the code that predicts age given lengths to make it more efficient and more based on the data. 

The predicted ages seem to be biased downwards compared to the age data available (e.g. predicted age of 5 compared to age given as 6). Should I try to correct for this bias? 

```{r cleaning, echo = FALSE, message = FALSE}
library(tidyverse)

### Read the data in
lingcod <- read.csv("data/LingcodwithGutContents_All_BB.csv")

## Calculate the sum of all the gut content in weight and create new column
## with total gut content in weight
lingcod_gut_content_wt <- lingcod %>% 
  select(Sample.ID, contains("wt.."))
lingcod_gut_content_wt[is.na(lingcod_gut_content_wt)] <- 0
lingcod_gut_content_wt <- lingcod_gut_content_wt %>% 
  mutate(wt..Total = rowSums(lingcod_gut_content_wt[,2:103])) %>% 
  select(Sample.ID, wt..Total)

## Calculate the sum of all the gut content by number
lingcod_gut_content_X <- lingcod %>% 
  select(Sample.ID, contains("X.."))
lingcod_gut_content_X[is.na(lingcod_gut_content_X)] <- 0
lingcod_gut_content_X <- lingcod_gut_content_X %>% 
  mutate(X..Total = rowSums(lingcod_gut_content_X[,2:103])) %>% 
  select(Sample.ID, X..Total)

# New working dataset with total food in stomach  
lingcod_working <- left_join(lingcod, lingcod_gut_content_wt, by = "Sample.ID")
lingcod_working <- left_join(lingcod_working, lingcod_gut_content_X, by = "Sample.ID")
lingcod_working <- lingcod_working %>% # Remove the NA rows at the end
  slice(1:1321)

# Create working dataset with relevant columns and make a
# new column grouping all Sebastes species into one
lingcod_rockfish <- lingcod_working %>%
  select(Sample.ID, Ages, Date..yymmdd., Port, Location, MPA.1.fish.take.prohibited., Lat, Long,
         Depth.ft, TL.cm, Sex.1, Wt.kg, Gape, Maturity, Stage, 
         Gut.wt.kg, Gut.contents.kg, X..Sebastes, wt..Sebastes, 
         X..Sebastes.jordani, wt..Sebastes.jordani, X..Yellowtail.Rockfish,
         wt..Yellowtail.Rockfish, X..Blue.Rockfish, wt..Blue.Rockfish,
         X..Black.Rockfish, wt..Black.RF, X..Puget.Sound.Rockfish,
         wt..Puget.Sound.Rockfish, wt..Total, X..Total) %>% 
  replace(is.na(.), 0) %>%
  mutate(X..Sebastes.Total = rowSums(select(.,X..Sebastes, X..Sebastes.jordani,
                                            X..Yellowtail.Rockfish, X..Blue.Rockfish,
                                            X..Black.Rockfish, X..Puget.Sound.Rockfish))) %>% 
  mutate(wt..Sebastes.Total = rowSums(select(.,wt..Sebastes, wt..Sebastes.jordani,
                                             wt..Yellowtail.Rockfish, wt..Blue.Rockfish,
                                             wt..Black.RF, wt..Puget.Sound.Rockfish))) %>% 
  filter(Sex.1 != "") ## We realized that rows with NA for Sex were all just 0s

# Replace all 0s in Ages column with NAs
lingcod_rockfish$Ages <- replace(lingcod_rockfish$Ages, lingcod_rockfish$Ages == 0, NA)

# Create a function that predicts age using the francis parameterization of the 
# VBGF. At ages that are too high, the function fails because it attempts to 
# take a log of a negative number which isn't possible. So, an if statement
# is used to set a specific age for lengths that are too large. This max age is 
# is determined by the max age in the observed data.

# MALES
ages_francis_male <- c(3,11) # male t1 and t3 age
male_TL_aget1 = lingcod_rockfish %>% # Filter out all male age at t1
  filter(Sex.1 == "M", Ages == ages_francis_male[1]) %>% 
  select(TL.cm)
male_TL_aget2 = lingcod_rockfish %>% # Filter out all male at t2
  filter(Sex.1 == "M", Ages == mean(ages_francis_male)) %>% 
  select(TL.cm)
male_TL_aget3 = lingcod_rockfish %>% # Filter out all male at t3
  filter(Sex.1 == "M", Ages == ages_francis_male[2]) %>% 
  select(TL.cm)
L1_m <- mean(male_TL_aget1$TL.cm) # mean length at relatively young age t1 (age 3 used here)
L2_m <- mean(male_TL_aget2$TL.cm) # mean length at the average of t1 and t2 (age 7 used here)
L3_m <- mean(male_TL_aget3$TL.cm) # mean length at relatively old age t2 (age 11 used here)
r_m <- (L3_m-L2_m)/(L2_m-L1_m)

predict_age_m <- function(length) { # Function that predicts male age given length
  if(length >= 83.8) {
    pred_age <-  14
  } else {
    pred_age <- ages_francis_male[1] + (ages_francis_male[2]-ages_francis_male[1])*((log(1-(1-r_m^2)*((length-L1_m)/(L3_m-L1_m))))/(2*log(r_m)))
  }
  return(pred_age) 
}

# FEMALES
ages_francis_female <- c(3,15)
female_TL_aget1 = lingcod_rockfish %>% # Filter out all female age at t1
  filter(Sex.1 == "F", Ages == ages_francis_female[1]) %>% 
  select(TL.cm)
female_TL_aget2 = lingcod_rockfish %>% # Filter out all female age at t2
  filter(Sex.1 == "F", Ages == mean(ages_francis_female)) %>% 
  select(TL.cm)
female_TL_aget3 = lingcod_rockfish %>% # Filter out all female age at t2
  filter(Sex.1 == "F", Ages == ages_francis_female[2]) %>% 
  select(TL.cm)
L1_f <- mean(female_TL_aget1$TL.cm)  # mean length at relatively young age t1 (age 3 used here)
L2_f <- mean(female_TL_aget2$TL.cm)  # mean length at the average of t1 and t2 (age 9 used here)
L3_f <- mean(female_TL_aget3$TL.cm) # mean length at relatively old age t2 (age 15 used here)
r_f <- (L3_f-L2_f)/(L2_f-L1_f)

predict_age_f <- function(length) {
  if(length >= 105) {
    pred_age <-  17
  } else {
    pred_age <- ages_francis_female[1] + (ages_francis_female[2]-ages_francis_female[1])*((log(1-(1-r_f^2)*((length-L1_f)/(L3_f-L1_f))))/(2*log(r_f)))
  }
  return(pred_age) 
}

# this helps prevent the warning messages.
v_predict_age_m <- Vectorize(predict_age_m)
v_predict_age_f <- Vectorize(predict_age_f)


# Add a new row with the predicted ages
## Males
lingcod_male <- lingcod_rockfish %>% 
  filter(Sex.1 == "M")

lingcod_male <- lingcod_male %>% 
  mutate(age_pred = v_predict_age_m(lingcod_male$TL.cm))

## Females

lingcod_female <- lingcod_rockfish %>% 
  filter(Sex.1 == "F")

lingcod_female <- lingcod_female %>% 
  mutate(age_pred = v_predict_age_f(lingcod_female$TL.cm))

### Let's remove the decimal places for the predicted age
lingcod_male$age_pred <- round(lingcod_male$age_pred)
lingcod_female$age_pred <- round(lingcod_female$age_pred)

### Create a new column with complete age (NA's filled in with predicted)
lingcod_male <- lingcod_male %>% 
  mutate(age_useful = Ages)

lingcod_female <- lingcod_female %>% 
  mutate(age_useful = Ages)

lingcod_male$age_useful[is.na(lingcod_male$age_useful)] <- lingcod_male$age_pred[is.na(lingcod_male$age_useful)]
lingcod_female$age_useful[is.na(lingcod_female$age_useful)] <- lingcod_female$age_pred[is.na(lingcod_female$age_useful)]

### Merge male and female dataset to re-create the lingcod_rockfish dataset
lingcod_rockfish <- rbind(lingcod_male, lingcod_female)

### Create a new column with the proportion of stomach content that is sebastes
lingcod_rockfish <- lingcod_rockfish %>% 
  mutate(gut.ratio.sebastes.wt = wt..Sebastes.Total/wt..Total) %>% 
  mutate(gut.ratio.sebastes.X = X..Sebastes.Total/X..Total)
lingcod_rockfish$gut.ratio.sebastes.wt[is.na(lingcod_rockfish$gut.ratio.sebastes.wt)] <- 0

# Create new dataset with just lingcod caught from California ports
CA_ports <- c("MOR", "BOD", "SDG", "FTB", "SLO", "LOS", "EUR", "HMB", "MON", "EME", "SBA")
lingcod_rockfish_CA <- lingcod_rockfish %>% 
  filter(Port %in% CA_ports)
```

To determine diet fraction, I created a new dataset "lingcod_rockfish_CA" which only looks at fish caught off ports within California. This reduced the number of observations from 1320 to 561. To calculate the fraction of lingcod diet that is rockfish, I simply divided the total Sebastes weight/number by the total weight/number of all gut contents for each lingcod individual. Below is a table with the average of these proportions for each lingcod age. There's quite a bit of variation to this number though, as there are a lot of 0s and several 1s in the diet fraction. 

```{r diet fraction, echo=FALSE, message = FALSE, warning = FALSE}
# Dataframe with the average diet fraction by weight that is sebastes for each age and sex 
dietfrac_by_age_wt <- lingcod_rockfish_CA %>% 
  group_by(Sex.1, age_useful) %>% 
  summarise(mean = mean(gut.ratio.sebastes.wt), sd = sd(gut.ratio.sebastes.wt), 
            n = n(), max = max(gut.ratio.sebastes.wt), min = min(gut.ratio.sebastes.wt))
dietfrac_by_age_wt$sd[is.na(dietfrac_by_age_wt$sd)] <- 0
dietfrac_by_age_wt

# Dataframe with the average diet fraction by number that is sebastes for each age and sex 
dietfrac_by_age_X <- lingcod_rockfish_CA %>% 
  group_by(Sex.1, age_useful) %>% 
  summarise(mean = mean(gut.ratio.sebastes.X), sd = sd(gut.ratio.sebastes.X), n = n())
dietfrac_by_age_X$sd[is.na(dietfrac_by_age_X$sd)] <- 0

# The proportion of stomachs that contained rockfish of all the stomachs
perc.occurrence <- (sum(lingcod_rockfish_CA$X..Sebastes.Total > 0)/nrow(lingcod_rockfish_CA))*100

### Proportion of gut content that is Sebastes across Ages
  # males
ggplot(lingcod_rockfish_CA[which(lingcod_rockfish_CA$Sex.1 == "M"),], aes(age_useful, gut.ratio.sebastes.wt)) +
  geom_point(color = "cyan3") +
  theme_classic() +
  labs( x = "Lingcod Age", y = "Fraction of diet that is Sebastes", title = "Male") +
  xlim(0,17)
  # females
ggplot(lingcod_rockfish_CA[which(lingcod_rockfish_CA$Sex.1 == "F"),], aes(age_useful, gut.ratio.sebastes.wt)) +
  geom_point(color = "lightcoral") +
  theme_classic() +
  labs( x = "Lingcod Age", y = "Fraction of diet that is Sebastes", title = "Female") +
  xlim(0,17)

#graphing the mean diet fraction by age
ggplot(dietfrac_by_age_wt, aes(age_useful, mean, col = Sex.1)) +
  geom_point() +
  labs( x = "Lingcod Age", y = "Average fraction of diet that is Sebastes") +
  theme_classic()
```


%O = 13% This is the proportion of stomachs that contained rockfish of all the stomachs  
%OF = 17.0% for females  
%OM = 10.7% for males  

How do I deal with ages that have 0 sebastes diet fraction (resulting from small sample size or even just bad timing). Should I try to fit a distribution to this data after removing the 0 in the older age classes? Should I bootstrap within age classes?

To add the lingcod consumption to the rockfish dynamic model, I calculate consumption as 

$Q =\sum_{s} \sum_{a_l=1}^{l} P_{a_l} L_{a_l,s}$ 

where $P$ is a vector of the proportion of lingcod diet that is composed of Sebastes by lingcod age.




